Compositional Certified Resource Bounds\ifshort{\vspace{-1.15cm}}{}

### 2014-12-11 paper/main.tex
attempt to explain the new increment rule

\iffull{
\begin{figure*}
\begin{mathpar}
%
\Rule[leftskip=.4cm,rightskip=.3cm]{Q:Skip}
{ }
{ B; R; (\Gamma, Q) \vdash \code{skip} \dashv (\Gamma, Q) }
%
\and
\Rule{Q:Break}
{ }
{ (\Gamma, Q_B); R; (\Gamma, Q_B) \vdash \code{break} \dashv (\Gamma', Q') }
%
\and
\Rule[leftskip=.3cm,rightskip=.4cm]{Q:Tick}
{ XXX }
{ B; R; (\Gamma, Q{+}n) \vdash \code{tick}(n) \dashv (\Gamma, Q) }
%
\and
\Rule{Q:Return}
{ P = Q_R[\Vret/x]
\\ \Gamma = \Gamma_R[\Vret/x]
\\ \forall i \in \dom{P} . \, p_i = q_i
}
{ B; (\Gamma_R,Q_R); (\Gamma, Q) \vdash \code{return}~x \dashv (\Gamma',Q') }
%
\and \Rule{Q:Update}
{ q'_{xy}, q'_{yx} \in \Q
\\ \forall u. (q_{yu} = q'_{xu} + q'_{yu} \land q_{uy} = q'_{ux} + q'_{uy})
}
{ B; R; (\Gamma[x/y], Q {+} M_u {+} M_e(y)) \vdash x \gets y \dashv (\Gamma, Q') }
%
\\ \Rule[leftskip=.1cm,rightskip=.1cm]{Q:Loop}
{ (\Gamma', Q'); R; (\Gamma,Q) \vdash S \dashv (\Gamma,  Q) }
{ B; R; (\Gamma, Q) \vdash \code{loop}~S \dashv (\Gamma', Q') }
%
\and \Rule[leftskip=.2cm,rightskip=.1cm]{Q:IncP}
{ \Gamma \models y \ge 0
\\ \textstyle q'_{0y} = q_{0y} - \sum_u \max(q_{ux}, -q_{xu})
}
{ \textstyle
  B; R; (\Gamma[x/x {+} y], Q) \vdash x \gets x + y \dashv (\Gamma, Q')
}
%
\and \Rule{Q:DecP}
{ \Gamma \models y \ge 0
\\ \textstyle q'_{y0} = q_{y0} - \sum_u \max(q_{xu}, -q_{ux})
}
{ \textstyle
  B; R; (\Gamma[x/x {-} y],Q) \vdash x \gets x - y \dashv (\Gamma, Q')
}
%
\and \Rule{Q:Inc}
{  \textstyle q'_{0y} = q_{0y} - \sum_u \max(q_{ux}, -q_{xu})
\\\\ \textstyle q'_{y0} = q_{y0} - \sum_u \max(q_{xu}, -q_{ux})
}
{ \textstyle
  B; R; (\Gamma[x/x {\pm} y], Q) \vdash x \gets x \pm y \dashv (\Gamma, Q')
}
%
\and \Rule{Q:If}
{ B; R; (\Gamma \land e, Q) \vdash S_1 \dashv (\Gamma',Q')
\\\\ B; R; (\Gamma \land \neg e, Q) \vdash S_2 \dashv (\Gamma',Q')
}
{ B; R; (\Gamma, Q{+}M_e(e)) \vdash \code{if}(e)~S_1~\code{else}~S_2 \dashv (\Gamma',Q') }
%
\and\Rule{Q:Seq}
{  B; R; (\Gamma, Q) \vdash S_1 \dashv (\Gamma', Q')
\\\\ B; R; (\Gamma', Q') \vdash S_2 \dashv (\Gamma'',Q'')
}
{ B; R; (\Gamma, Q) \vdash S_1;S_2 \dashv (\Gamma'',Q'') }
%
\and\Rule{Q:Call}
{ (\Gamma_f, Q_f,  \Gamma_f', Q_f') \in \Delta(f) \!
\\ \text{Loc} = \text{Locals}(Q) \!
\\ \forall i \neq j . \, x_i \neq x_j \!
\\ c \in \Qplusz \!
\\ Q = P + S \!
\\ Q' = P' + S
\\ U = Q_f[\Vargs / \vec x]
\\ U' = Q'_f[\Vret/r]
\\ \forall i \in \dom{U} . \, p_i = u_i
\\ \forall i \in \dom{U'} .\, p'_i = u'_i
\\ \forall i \not\in \dom{U'} .\, p'_i = 0
\\ \forall i \not\in \text{Loc}
  .\, s_i = 0
}
{ B; R; (\Gamma_f[\Vargs / \vec x] \land \Gamma_{\text{Loc}}, Q {+} c)
  \vdash r \gets f(\vec x) \dashv
  (\Gamma_f'[\Vret/r] \land \Gamma_{\text{Loc}}, Q' {+} c)
}
%
\and
\Rule[leftskip=.2cm,rightskip=.2cm]{Q:Assert}
{ }
{ B; R; (\Gamma, Q) \vdash \code{assert}~e \dashv (\Gamma \land e, Q) }
%
\and
\Rule[leftskip=.2cm,rightskip=.2cm]{Q:Extend}
{ \fenv f = (\vec y, S_f)
\\\\ B; (\Gamma_f', Q'_f); (\Gamma_f[\Vargs/\vec y], Q_f[\Vargs/ \vec y] ) \vdash S_f \dashv (\Gamma', Q')
}
{ (\Gamma_f, Q_f, \Gamma_f', Q_f') \in \Delta(f) }
%
\and \Rule{Q:Weak}
{ B; R; (\Gamma_2,Q_2) \vdash S \dashv (\Gamma'_2, Q'_2)
\\ \!\! \Gamma_1 \models \Gamma_2
\\\\ Q_1 \succeq_{\Gamma_1} Q_2
\\ \Gamma'_2 \models \Gamma'_1
\\ Q'_2 \succeq_{\Gamma_2'} Q'_1
}
{ B; R; (\Gamma_1,Q_1) \vdash S \dashv (\Gamma'_1, Q'_1) }
%
\and \Rule{Relax}
{ \mathcal L = \{ xy \mid \exists l_{xy} {\in} \N \, . \, \Gamma \models l_{xy} \le |\inter x y| \}
\\ \mathcal U = \{ xy \mid \exists u_{xy}{\in} \N \, . \, \Gamma \models |\inter x y| \le u_{xy} \}
\\\\ \forall i \in \mathcal U.\, q'_i \ge q_i - r_i
\\ \forall i \in \mathcal L.\, q'_i \ge q_i + p_i
\\ \forall i \not\in \mathcal U {\cup} \mathcal L {\cup} \{0\} .\, q'_i \ge q_i
\\ q'_0 \geq q_0 {\textstyle + \sum_{i \in \mathcal U} u_i r_i - \sum_{i \in \mathcal L} l_i p_i}
}
{ Q' \succeq_\Gamma Q }
\end{mathpar}
\caption{Inference rules of the quantitative analysis.}
\label{fig:auto}
\end{figure*}
}{}

\ifshort{
\begin{figure*}
\small
\vspace{-.4cm}
\def \MathparLineskip {\lineskip=0.35cm}
\begin{mathpar}
\Rule[leftskip=.4cm,rightskip=.3cm]{Q:Skip}
{ }
{ \!\! B; R; (\Gamma, Q) \vdash \code{skip} \dashv (\Gamma, Q) \!\!}
%
\and \Rule[leftskip=.1cm,rightskip=.1cm]{Q:Loop}
{ (\Gamma', Q'); R; (\Gamma,Q) \vdash S \dashv (\Gamma,  Q {+} M_l) }
{ B; R; (\Gamma, Q) \vdash \code{loop}~S \dashv (\Gamma', Q') }
%
\\ \Rule[leftskip=1cm,rightskip=.5cm]{Q:IncP}
{ \Gamma \models y \ge 0 \!
\\ \mathcal U = \left\{ u \mid \Gamma \models x + y \in \inter x u \right\} \!
\\\\ \textstyle q'_{0y} = q_{0y}
      + \sum_{u \in \mathcal U} q_{xu}
      - \sum_{v \not\in \mathcal U} q_{vx} \!
}
{ \textstyle
  B; R; (\Gamma[x/x {+} y], Q {+} M_u {+} M_e(x {+} y)) \vdash x \gets x + y \dashv (\Gamma, Q')
}
%
\and \Rule[leftskip=.5cm,rightskip=1cm]{Q:Inc}
{ M = M_u + M_e(x {\pm} y)
\\\\ \textstyle q'_{0y} = q_{0y} - \sum_{v} q_{vx}
\\ \textstyle q'_{y0} = q_{y0} - \sum_{v} q_{xv}
}
{ \textstyle
  B; R; (\Gamma[x/x {\pm} y], Q {+} M) \vdash x \gets x \pm y \dashv (\Gamma, Q')
}
%
\\ \Rule{Q:Seq}
{  B; R; (\Gamma, Q) \vdash S_1 \dashv (\Gamma', Q' {+} M_s)
\\\\ B; R; (\Gamma', Q') \vdash S_2 \dashv (\Gamma'',Q'')
}
{ B; R; (\Gamma, Q) \vdash S_1;S_2 \dashv (\Gamma'',Q'') }
%
\and \Rule{Q:Weak}
{ B; R; (\Gamma_2,Q_2) \vdash S \dashv (\Gamma'_2, Q'_2)
\\ \!\! \Gamma_1 \models \Gamma_2
\\\\ Q_1 \succeq_{\Gamma_1} Q_2
\\ \Gamma'_2 \models \Gamma'_1
\\ Q'_2 \succeq_{\Gamma_2'} Q'_1
}
{ B; R; (\Gamma_1,Q_1) \vdash S \dashv (\Gamma'_1, Q'_1) }
%
\\ \Rule{Relax}
{ \mathcal L = \{ xy \mid \exists l_{xy} {\in} \N \, . \, \Gamma \models l_{xy} \le |\inter x y| \}
\\ \mathcal U = \{ xy \mid \exists u_{xy}{\in} \N \, . \, \Gamma \models |\inter x y| \le u_{xy} \}
\\\\ \forall i \in \mathcal U.\, q'_i \ge q_i - r_i
\\ \forall i \in \mathcal L.\, q'_i \ge q_i + p_i
\\ \forall i \not\in \mathcal U {\cup} \mathcal L {\cup} \{0\} .\, q'_i \ge q_i
\\ q'_0 \geq q_0 {\textstyle + \sum_{i \in \mathcal U} u_i r_i - \sum_{i \in \mathcal L} l_i p_i}
}
{ Q' \succeq_\Gamma Q }
\vspace{-.3cm}
\end{mathpar}
\caption{Selected inference rules of the quantitative analysis.}
\label{fig:auto}
\end{figure*}
}{}

@@

The derivation system for the automatic amortized analysis is defined
in~\pref{fig:auto}.  The derivation rules derive judgements of the form
$$
  (\Gamma_B, Q_B); (\Gamma_R, Q_R);
    (\Gamma, Q) \vdash S \dashv (\Gamma', Q').
$$
%
The part $(\Gamma, Q) \vdash \!\! S \!\! \dashv (\Gamma', Q')$ of the judgement
can be seen as a quantitative Hoare triple.  All assertions are split into
two parts, the logical part and the quantitative part.  The
quantitative part $Q$ represents a potential function as a collection
of non-negative numbers $q_i$ indexed by the index set $I$.  The
logical part $\Gamma$ is left abstract but is enforced by our
derivation system to respect classic Hoare logic constraints.
%
The meaning of this basic judgment is as follows: If $S$ is executed
with starting state $\state$, the assertions in $\Gamma$
hold, and at least $Q(\state)$ resources are available then the
evaluation does not run out of resources and, if the execution
terminates in state $\state'$, there are at least $Q(\state')$
\mbox{resources left and $\Gamma'$ holds for $\state'$.}

@@

As a convention, if $Q$ and $Q'$ are quantitative annotations
we assume that $Q = (q_i)_{i\in I}$ and $Q' = (q'_i)_{i \in I}$.
The notation $Q \pm n$ used in many rules defines a new context $Q'$
such that $q'_0 = q_0 \pm n$ and $\forall i \neq 0 .\, q'_i = q_i$.
In all the rules, we have the implicit side condition that all rational
coefficients are non-negative.  Finally, if a rule mentions $Q$ and $Q'$
and leaves the latter undefined at some index $i$ we assume that $q'_i = q_i$.

We describe the automatic amortized analysis for a subset of
expressions of Clight.  Assignments are restricted to the form $x \gets
y$ or $x \gets x \pm y$.  In the implementation, a Clight
program is converted into this form prior to analysis without changing
the resource cost.  This is achieved by using a series of \emph{cost-free
  assignments} that do not result in additional cost in the
semantics.  Non-linear operations such as $x \gets z*y$ or $x \gets
a[y]$ are handled by assigning $0$ to coefficients
like $q_{xa}$ and $q_{ax}$ that contain $x$ after the assignment.

@@

During the analysis, function specifications are quadruples
$(\Gamma_f, Q_f, \Gamma'_f, Q'_f)$ where $\Gamma_f, Q_f$ depend on
$\Vargs$, and $\Gamma'_f, Q'_f$ depend on $\Vret$.  These parameters
are instantiated by appropriate variables on call sites.  A distinctive
feature of our analysis is that it respects the function abstraction:
when deriving a function specification it generates a set of
constraints and the above quadruple; once done, the constraint
set can readily be reused for every call site and the function need
not be analyzed multiple times.
%
Therefore, the derivation rules are parametric in a function context $\Delta$
that we leave implicit in the rules presented here. More details
can be found in the extended version.

@@

The rules of our derivation system must serve two purposes.  They must
\emph{attach} potential to certain program variable intervals and use
this potential, when it is allowed, to \emph{pay} for resource consuming
operations.  These two purposes are illustrated on the {\sc Q:Skip}
rule.  This rule reuses its precondition as postcondition, it is
explained by two facts: first, no resource is consumed by the skip
operation, thus no potential has to be used to pay for the evaluation;
second, the program state is not changed by
the execution of a skip statement, thus all the potential available
before the execution of the skip statement is still available after.
\iffull{
These two observations would also justify a rule with the side-condition
$Q \ge Q'$, because some potential is always safely lost.  For the sake
of orthogonality, it is not how we present the rule and this form of
weakening is in fact expressed as the stand-alone rule {\sc Q:Weak}
described later.}{}

@@

The rules {\sc Q:IncP}\iffull{, {\sc Q:DecP},}{} and {\sc Q:Inc}
describe how the potential is distributed after a size change of
a variable.
\iffull{The rule {\sc Q:IncP} is for increments $x \gets x + y$
and {\sc Q:DecP} is for decrements $x \gets x - y$, they both apply only
when we can deduce from the logical context $\Gamma$ that $y \ge 0$.
Of course, we have symmetrical rules {\sc Q:IncN} and {\sc Q:DecN}
that can be applied if $y$ is not positive.}{The rule {\sc Q:IncP} is for increments
$x \gets x+y$, it applies when we can deduce from the logical context
$\Gamma$ that $y \ge 0$.  Of course, we have symmetrical rules that
can be applied for decrements and when $y$ is not positive.}
The rules are all equivalent in the case where $y=0$.
The rule {\sc Q:Inc} can be applied
if we cannot deduce any information about $y$.  In $\toolname$,
the rules for increments and decrements are combined into one rule\iffull{ where the specialized rules take priority over {\sc Q:Inc}.}{.}

\iffull{To explain how rules for increment and decrement work, it is
sufficient to understand the rule {\sc Q:IncP}.  The others follow the
same idea and are symmetrical.}{}  In {\sc Q:IncP}, the program updates a
variable $x$ with $x+y$ where $y \ge 0$.  Since $x$ is changed, the
quantitative annotation must be updated to reflect the change of the
program state.  We write $x'$ for the value of $x$ after the assignment.
Since $x$ is the only variable changed, only intervals of the form
$\inter u x$ and $\inter x u$ will be resized. Note that for any $u$,
$\inter x u$ will get smaller with the update, and if $x' \in \inter x
u$ we have $|\inter x u| = |\inter x {x'}| + |\inter {x'} u|$.  But
$|\inter x {x'}| = |\inter 0 y|$ which means that the potential
$q'_{0y}$ in the postcondition can be increased by $q_{xu}$ under the
guard that $x' \in \inter x u$.  Dually, the interval $\inter v x$ can
get bigger with the update.  We know that $|\inter v {x'}| \le y +
|\inter v x|$.  So we decrease the potential of $\inter 0 y$ by
$q_{vx}$ to pay for this change.  The rule ensures this
only for $v \not\in \mathcal U$ because $x \le v$
otherwise, and thus $|\inter v x| = 0$.

>>>>

\iffull{
\begin{figure*}
\begin{mathpar}
%
\Rule[leftskip=.4cm,rightskip=.3cm]{Q:Skip}
{ }
{ B; R; (\Gamma, Q) \vdash \code{skip} \dashv (\Gamma, Q) }
%
\and
\Rule{Q:Break}
{ }
{ (\Gamma, Q_B); R; (\Gamma, Q_B) \vdash \code{break} \dashv (\Gamma', Q') }
%
\and
\Rule[leftskip=.3cm,rightskip=.4cm]{Q:Tick}
{ n < 0 \implies Q \ge 0 }
{ B; R; (\Gamma, Q) \vdash \code{tick}(n) \dashv (\Gamma, Q{-}n) }
%
\and
\Rule{Q:Return}
{ P = Q_R[\Vret/x]
\\ \Gamma = \Gamma_R[\Vret/x]
\\ \forall i \in \dom{P} . \, p_i = q_i
}
{ B; (\Gamma_R,Q_R); (\Gamma, Q) \vdash \code{return}~x \dashv (\Gamma',Q') }
%
\and \Rule{Q:Update}
{ q'_{xy}, q'_{yx} \in \Q
\\ \forall u. (q_{yu} = q'_{xu} + q'_{yu} \land q_{uy} = q'_{ux} + q'_{uy})
}
{ B; R; (\Gamma[x/y], Q {+} M_u {+} M_e(y)) \vdash x \gets y \dashv (\Gamma, Q') }
%
\and \Rule[leftskip=.1cm,rightskip=.1cm]{Q:Loop}
{ (\Gamma', Q'); R; (\Gamma,Q) \vdash S \dashv (\Gamma,  Q) }
{ B; R; (\Gamma, Q) \vdash \code{loop}~S \dashv (\Gamma', Q') }
%
\and \Rule{Q:Inc}
{  \textstyle q'_{0y} = q_{0y} - \sum_u \max(q_{ux}, -q_{xu})
\\\\ \textstyle q'_{y0} = q_{y0} - \sum_u \max(q_{xu}, -q_{ux})
}
{ \textstyle
  B; R; (\Gamma[x/x {+} y], Q) \vdash x \gets x + y \dashv (\Gamma, Q')
}
%
\and \Rule{Q:Dec}
{  \textstyle q'_{0y} = q_{0y} - \sum_u \max(q_{xu}, -q_{ux})
\\\\ \textstyle q'_{y0} = q_{y0} - \sum_u \max(q_{ux}, -q_{xu})
}
{ \textstyle
  B; R; (\Gamma[x/x {-} y], Q) \vdash x \gets x - y \dashv (\Gamma, Q')
}
%
\and \Rule{Q:If}
{ B; R; (\Gamma \land e, Q) \vdash S_1 \dashv (\Gamma',Q')
\\\\ B; R; (\Gamma \land \neg e, Q) \vdash S_2 \dashv (\Gamma',Q')
}
{ B; R; (\Gamma, Q) \vdash \code{if}(e)~S_1~\code{else}~S_2 \dashv (\Gamma',Q') }
%
\and\Rule{Q:Seq}
{  B; R; (\Gamma, Q) \vdash S_1 \dashv (\Gamma', Q')
\\\\ B; R; (\Gamma', Q') \vdash S_2 \dashv (\Gamma'',Q'')
}
{ B; R; (\Gamma, Q) \vdash S_1;S_2 \dashv (\Gamma'',Q'') }
%
\and\Rule{Q:Call}
{ (\Gamma_f, Q_f,  \Gamma_f', Q_f') \in \Delta(f) \!
\\ \text{Loc} = \text{Locals}(Q) \!
\\ \forall i \neq j . \, x_i \neq x_j \!
\\ c \in \Qplusz \!
\\ Q = P + S \!
\\ Q' = P' + S
\\ U = Q_f[\Vargs / \vec x]
\\ U' = Q'_f[\Vret/r]
\\ \forall i \in \dom{U} . \, p_i = u_i
\\ \forall i \in \dom{U'} .\, p'_i = u'_i
\\ \forall i \not\in \dom{U'} .\, p'_i = 0
\\ \forall i \not\in \text{Loc}
  .\, s_i = 0
}
{ B; R; (\Gamma_f[\Vargs / \vec x] \land \Gamma_{\text{Loc}}, Q {+} c)
  \vdash r \gets f(\vec x) \dashv
  (\Gamma_f'[\Vret/r] \land \Gamma_{\text{Loc}}, Q' {+} c)
}
%
\and \Rule[leftskip=.2cm,rightskip=.2cm]{Q:Assert}
{ }
{ B; R; (\Gamma, Q) \vdash \code{assert}~e \dashv (\Gamma \land e, Q) }
%
\and \Rule[leftskip=.2cm,rightskip=.2cm]{Q:Extend}
{ \fenv f = (\vec y, S_f)
\\ Q' \ge 0
\\ Q'_f \ge 0
\\\\ B; (\Gamma_f', Q'_f); (\Gamma_f[\Vargs/\vec y], Q_f[\Vargs/ \vec y] ) \vdash S_f \dashv (\Gamma', Q')
}
{ (\Gamma_f, Q_f, \Gamma_f', Q_f') \in \Delta(f) }
%
\and \Rule{Q:Weak}
{  \!\! \Gamma_1 \models \Gamma_2
\\ Q_1 \succeq_{\Gamma_1} Q_2
\\ B; R; (\Gamma_2,Q_2) \vdash S \dashv (\Gamma'_2, Q'_2)
\\ \Gamma'_2 \models \Gamma'_1
\\ Q'_2 \succeq_{\Gamma_2'} Q'_1
}
{ B; R; (\Gamma_1,Q_1) \vdash S \dashv (\Gamma'_1, Q'_1) }
%
\and \Rule{Relax}
{ \mathcal L = \{ xy \mid \exists l_{xy} {\in} \N \, . \, \Gamma \models l_{xy} \le |\inter x y| \}
\\ \mathcal U = \{ xy \mid \exists u_{xy}{\in} \N \, . \, \Gamma \models |\inter x y| \le u_{xy} \}
\\ \forall i.\, p_i, r_i \in \Qplusz
\\\\ \forall i \in \mathcal U.\, q'_i \ge q_i - r_i
\\ \forall i \in \mathcal L.\, q'_i \ge q_i + p_i
\\ \forall i \not\in \mathcal U {\cup} \mathcal L {\cup} \{0\} .\, q'_i \ge q_i
\\ q'_0 \geq q_0 {\textstyle + \sum_{i \in \mathcal U} u_i r_i - \sum_{i \in \mathcal L} l_i p_i}
}
{ Q' \succeq_\Gamma Q }
\end{mathpar}
\caption{Inference rules of the quantitative analysis.}
\label{fig:auto}
\end{figure*}
}{}

\ifshort{
                    \jan{XXX PLEASE INSERT FIGURE HERE XXX}
}{}

@@

The derivation system for the automatic amortized analysis is defined
in~\pref{fig:auto}.  The derivation rules yield judgements of the form
$$
  (\Gamma_B, Q_B); (\Gamma_R, Q_R);
    (\Gamma, Q) \vdash S \dashv (\Gamma', Q').
$$
%
The part $(\Gamma, Q) \! \vdash \! S \! \dashv \! (\Gamma', Q')$ of the judgement
can be seen as a quantitative Hoare triple.  All assertions are split into
two parts, the logical part and the quantitative part.  The
quantitative part $Q$ represents a potential function as a collection
of rational numbers $q_i$ indexed by the index set $I$.  The
logical part $\Gamma$ is left abstract but is enforced by our
derivation system to respect classic Hoare logic constraints.
%
The meaning of this basic judgment is as follows: If $S$ is executed
with starting state $\state$, the assertions in $\Gamma$
hold, $Q' {\ge} 0$, and at least $Q(\state)$ resources are available then the
evaluation does not run out of resources and, if the execution
terminates in state $\state'$, there are at least $Q'(\state')$
resources left and $\Gamma'$ holds for $\state'$.

@@

As a convention, if $Q$ and $Q'$ are quantitative annotations
we assume that $Q = (q_i)_{i\in I}$ and $Q' = (q'_i)_{i \in I}$.
The notation $Q \pm n$ used in many rules defines a new context $Q'$
such that $q'_0 = q_0 \pm n$ and $\forall i \neq 0 .\, q'_i = q_i$.
%
% THE FOLLOWING IS NO LONGER TRUE
% In all the rules, we have the implicit side condition that all rational
% coefficients are non-negative.
%
Finally, if a rule mentions $Q$ and $Q'$
and leaves the latter undefined at some index $i$ we assume that $q'_i = q_i$.

We describe the automatic amortized analysis for a subset of
expressions of Clight.  Assignments are restricted to the form $x \gets
y$ or $x \gets x \pm y$.  In the implementation, a Clight
program is converted into this form prior to analysis without changing
the resource cost.  This is achieved by using a series of \emph{cost-free
  assignments} that do not result in additional cost in the
semantics.  Non-linear operations such as $x \gets z*y$ or $x \gets
a[y]$ are handled by assigning $0$ to coefficients
like $q_{xa}$ and $q_{ax}$ that depend on $x$ after the assignment.

@@

During the analysis, function specifications are quadruples
$(\Gamma_f, Q_f, \Gamma'_f, Q'_f)$ where $\Gamma_f, Q_f$ depend on
$\Vargs$, and $\Gamma'_f, Q'_f$ depend on $\Vret$.  These parameters
are instantiated by appropriate variables on call sites.  A distinctive
feature of our analysis is that it respects the function abstraction:
when deriving a function specification it generates a set of
constraints and the above quadruple; once done, the constraint
set can readily be reused for every call site and the function need
not be analyzed multiple times.
%
Therefore, the derivation rules are parametric in a function context $\Delta$
that we leave implicit in the rules presented here.
% More details can be found in the extended version.

@@

The rules of our derivation system must serve two purposes.  They must
\emph{attach} potential to certain program variable intervals and use
this potential, when it is allowed, to \emph{pay} for resource consuming
operations.  These two purposes are illustrated on the {\sc Q:Skip}
rule.  This rule reuses its precondition as postcondition, it is
explained by two observations: first, no resource is consumed by the skip
operation, thus no potential has to be used to pay for the evaluation;
second, the program state is not changed by
the execution of a skip statement, thus all the potential available
before the execution of the skip statement is still available after.
\iffull{
These two observations would also justify a rule with the side-condition
$Q \ge Q'$, because some potential is always safely lost.  For the sake
of orthogonality, it is not how we present the rule and this form of
weakening is in fact expressed as the stand-alone rule {\sc Q:Weak}
described later.}{}

@@

The rules {\sc Q:Inc} and {\sc Q:Dec} describe how the potential is
distributed after a size change of a variable.
The rule {\sc Q:Inc} is for incements $x \gets x+y$ and the symmetric
rule {\sc Q:Dec} is for decrements $x \gets x-y$.
The two rules are equivalent in the case where $y=0$.

The two previous rules are essentially the same, so we only focus on
the explanation of {\sc Q:Inc}. In this rule, the program updates a
variable $x$ with $x+y$.  Since $x$ is changed, the
quantitative annotation must be updated to reflect the change of the
program state.  We write $x'$ for the value of $x$ after the assignment.
Since $x$ is the only variable changed, only intervals of the form
$\inter u x$ and $\inter x u$ will be resized.  Consider $S(x) = q_{xu}
|\inter x u| + q_{ux} |\inter u x|$, during the increment, $S$ will
increase by at most $y \cdot \max(q_{ux}, -q_{xu})$.
This means that we must decrease the potential $q_{0y}$ of $|\inter 0 y|$
by $\max(q_{ux}, -q_{xu})$ to pay for the change on
$|\inter x u|$ and $|\inter u x|$.  A symmetric argument explains the
change on $q_{y0}$.
